\documentclass{article}


\usepackage[margin=0.6in]{geometry}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{placeins}
\usepackage{nicefrac}
\usepackage{enumerate}
\usepackage{cancel}
\usepackage{array}
\usepackage{color}
\newcommand{\Rl}{\mathbb{R}}
\newcommand{\qiq}{\ \ \ \implies\ \ \ }
\newcommand{\vari}[1]{\text{var}\qty[#1]}
\newcommand{\cor}[1]{\text{cor}\qty[#1]}
\newcommand{\expec}[1]{\mathbb{E}\qty[#1]}
\newcommand{\cov}[1]{\text{cov}\qty[#1]}
\newcommand{\f}[3]{#1\ :\ #2 \rightarrow #3}
\newcommand{\prob}[1]{\mathbb{P}\qty[#1]}
\newcommand{\half}{\nicefrac{1}{2}}

\title{PBG 200A Notes}
\author{Sam Fleischer}
\date{November 21, 2016}

\begin{document}
    \maketitle

    \begin{itemize}
        \item Suppose $ab$ and $AB$ are more advantageous than $aB$ and $Ab$, we can select against recombination between the loci through inversion.
        \item A species having sex is \emph{not} the same as a species having different sexes.
        \item Males are defined through small mobile gametes and females are defined through larger less mobile gametes.
        \item Separate sexes has arisen multiple times as a result of sex.
        \item Mammals have XY, birds have ZW.
        \begin{itemize}
            \item in XY, male heterogamy
            \item in ZW, female heterogamy
        \end{itemize}
        \item X and Y evolved from an ancestral autosome - used to be the same size but at some point a sex-determining gene arose.
        \item Heteromorphic sex chromosomes have evolved multiple times, independently.
        \item If there is a male beneficial allele near a male determining allele, we can lock them together in an inversion.
        \item Then that section of the Y chromosome no longer recombines.  It's always homozygous.
        \item Deleterious alleles can be swept by the inversion and there is no way for the Y chromosome to get rid of them through recombination.
        \item More beneficial alleles and inversion links more loci together.
        \item Allopatric speciation
        \begin{itemize}
            \item two pops split due to geography
            \item low rate of migration causes independent genetic drift
            \item barriers to reproduction come as a passive side effect of anagenesis
            \item may happen rapidly if ecology of populations differ
        \end{itemize}
        \item Sympatric speciation
        \begin{itemize}
            \item Need strong selection on phenotypic extremes
            \item Need individuals to mate assortatively
            \item This is helped if assortative mating gene is close to the gene determining selected phenotype
            \item 
        \end{itemize}
    \end{itemize}

\end{document}















